Online Appendix

### 2024-07-29 draft/ms-adj-appendix.Rmd
replications

---
output:
  # bookdown::word_document2:
    # reference_docx: "word-template.docx"
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
    number_sections: no
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #latex_engine: xelatex
always_allow_html: true
header-includes:
  #- \usepackage{setspace}\doublespace
  # - \usepackage[nolists, fighead, tabhead]{endfloat}
  # - \usepackage{endnotes}
  # - \let\footnote=\endnote
  # - \setlength{\headheight}{14.5pt}
  # - \setlength{\headheight}{13.6pt}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \lhead{N.I. Hoffmann}
  - \rhead{`r format(Sys.time(), '%B %e, %Y')`}
# - \newcommand{\beginsupplement}{\setcounter{table}{0}  
# \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} 
# \renewcommand{\thefigure}{A\arabic{figure}}}

@@

title: "Double Robust, Flexible Adjustment Methods for Causal Inference: An Overview and an Evaluation"

@@

thanks: "I deeply thank  my committee -- Chad Hazlett, Onyebuchi A. Arah, Jennie E. Brand -- for their extensive suggestions and advice on previous drafts. I would also like to thank the members of UCLA's Practical Causal Inference Lab, UCLA's Social Inequality Data Science Lab, participants of the 2023 American Sociological Association's Annual Meeting, and participants of the 2024 All-UC Demography Conference for their feedback on previous versions of this paper."

abstract: "Double robust methods for flexible covariate adjustment in causal inference have proliferated in recent years. Despite their apparent advantages, these methods remain underutilized by social scientists. It is also unclear whether these methods actually outperform more traditional methods in finite samples. This paper has two aims: It is a guide to some of the latest methods in double robust, flexible covariate adjustment for causal inference, and it compares these methods to more traditional statistical methods and flexible \"single robust\" methods. It does this by using both simulated data where the treatment effect estimate is known, and then using comparisons of experimental and observational data from the National Supported Work Demonstration. Methods covered include Augmented Inverse Probability Weighting (AIPW), Targeted Maximum Likelihood Estimation (TMLE), and Double/Debiased Machine Learning (DML). Results suggest that some of these methods do outperform traditional methods in a wide range of simulations, but only slightly. In particular, the top performers are TMLE and AIPW in conjunction with flexible machine learning algorithms. But G-computation with the same flexible machine learning algorithms obtains almost identical results, and simple regression methods are nearly comparable in bias and are much more computationally efficient."

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = T, dev = c('pdf', 'png'), fig.retina = 3)
options("yaml.eval.expr" = TRUE)

@@

perform <- function(est_df, label){
  est_df %>%
    # mutate(d = d/truth,
    #        truth = 1) %>%
    summarize(bias = mean(d - truth),
              percent_bias = bias/sd(d),
              rmse = sqrt(mean((ate - truth)^2)),
              mae = median(abs(d - truth))
              # fail_count = first(fail_count),
              # comp_time = first(comp_time)
              ) %>%
    mutate(label = label,
           n = nrow(est_df) + fail_count) %>%
    select(label, everything()) %>%
    return()
}
```

```{r load-sims, eval = F}
lalonde_bootstrap <- function(dataset, sample, include_74 = F, iter = 100){

  lalonde_variables <- c('age', 'education', 'black', 'hispanic', 'married', 'nodegree', 're75', 're75_0')
  if(include_74 == F){
        dataset <- bind_rows(filter(lalonde_exp, treat == 1), dataset)

      } else {
        lalonde_variables <- c(lalonde_variables, 're74', 're74_0')
        dataset <- bind_rows(filter(lalonde_exp_74, treat == 1), dataset)
      }

  bootstrap_list <- list()

  for(i in 1:iter){
    dataset_bootstrap <- sample_n(dataset, nrow(dataset), replace = T)

    bootstrap_list[[i]] <- list(y = dataset_bootstrap$re78,
                  z = ifelse(dataset_bootstrap$treat == 1, 'trt', 'ctl'),
                  x = dataset_bootstrap[,lalonde_variables],
                  y.1 = NA, y.0 = NA,
                  dataset = i,
                  set = sample,
                  size = nrow(dataset_bootstrap))
  }

  return(bootstrap_list)
}

lalonde_setup <- function(dataset, sample, include_74 = F){

  lalonde_variables <- c('age', 'education', 'black', 'hispanic', 'married', 'nodegree', 're75', 're75_0')
  if(include_74 == F){
        dataset <- bind_rows(filter(lalonde_exp, treat == 1), dataset)

      } else {
        lalonde_variables <- c(lalonde_variables, 're74', 're74_0')
        dataset <- bind_rows(filter(lalonde_exp_74, treat == 1), dataset)
      }

    list(y = dataset$re78,
                  z = ifelse(dataset$treat == 1, 'trt', 'ctl'),
                  x = dataset[,lalonde_variables],
                  y.1 = NA, y.0 = NA,
                  dataset = i,
                  set = sample,
                  size = nrow(dataset)) %>%
      return()
}

set.seed(1859)
sims_lalonde <- c(
  lalonde_bootstrap(filter(lalonde_exp, treat == 0), sample = 'lalonde experimental original'),
  lalonde_bootstrap(lalonde_psid1_controls, 'lalonde PSID-1 original'),
  lalonde_bootstrap(lalonde_psid3_controls, 'lalonde PSID-3 original'),
  lalonde_bootstrap(lalonde_cps1_controls, 'lalonde CPS-1 original'),
  lalonde_bootstrap(lalonde_cps1_controls, 'lalonde CPS-3 original'),
  lalonde_bootstrap(filter(lalonde_exp_74, treat == 0), sample = 'lalonde experimental 74', include_74 = T),
  lalonde_bootstrap(lalonde_psid1_controls, 'lalonde PSID-1 74', include_74 = T),
  lalonde_bootstrap(lalonde_psid3_controls, 'lalonde PSID-3 74', include_74 = T),
  lalonde_bootstrap(lalonde_cps1_controls, 'lalonde CPS-1 74', include_74 = T),
  lalonde_bootstrap(lalonde_cps1_controls, 'lalonde CPS-3 74', include_74 = T))

# sims <- readRDS(here('files', 'sims.RDS'))

sims_main <- readRDS(here('files_main', 'sims.RDS'))
sims_large <- readRDS(here('files_large', 'sims_large.RDS'))
sims_small <- readRDS(here('files_small', 'sims_small.RDS'))

sims <- sims_main %>% append(sims_small) %>% append(sims_large) %>% append(sims_lalonde)

## Full lalonde
# sims <- list(
#   lalonde_setup(filter(lalonde_exp, treat == 0), sample = 'lalonde experimental original'),
#   lalonde_setup(lalonde_psid1_controls, 'lalonde PSID-1 original'),
#   lalonde_setup(lalonde_psid3_controls, 'lalonde PSID-3 original'),
#   lalonde_setup(lalonde_cps1_controls, 'lalonde CPS-1 original'),
#   lalonde_setup(lalonde_cps1_controls, 'lalonde CPS-3 original'),
#   lalonde_setup(filter(lalonde_exp_74, treat == 0), sample = 'lalonde experimental 74', include_74 = T),
#   lalonde_setup(lalonde_psid1_controls, 'lalonde PSID-1 74', include_74 = T),
#   lalonde_setup(lalonde_psid3_controls, 'lalonde PSID-3 74', include_74 = T),
#   lalonde_setup(lalonde_cps1_controls, 'lalonde CPS-1 74', include_74 = T),
#   lalonde_setup(lalonde_cps1_controls, 'lalonde CPS-3 74', include_74 = T))

# sims <- sims %>% append(sims_large)

# sims <- readRDS(here('files', 'sims_small.RDS'))

# for(i in 1:length(sims_small)){
#   # sims_large[[i]]$size <- nrow(sims_large[[i]]$x)
#   sims_small[[i]]$dataset <- 7
# }
# saveRDS(sims_large, here('files_large', 'sims_large.RDS'))
# 
# 
# sims_combined <- sims %>% append(sims_small) %>% append(sims_large)
# 
# index_df <- data.frame(dorie_dataset = sapply(sims_combined, function(x) x$dataset),
#                        set = sapply(sims_combined, function(x) x$set),
#                        size = sapply(sims_combined, function(x) x$size)) %>%
#   mutate(dataset = c(1:length(sims), 1:length(sims_small), 1:length(sims_large)))
# 
# lm_df <- read_csv(here('files_main', 'lm_df.csv')) %>% 
#   mutate(set = c(rep('main', 200), rep('linear', 200))) %>%
#   bind_rows(read_csv(here('files_small', 'lm_df.csv')) %>% mutate(set = 'small')) %>%
#   bind_rows(read_csv(here('files_large', 'lm_df.csv')) %>% mutate(set = 'large')) %>%
#   left_join(index_df)
# 
# 
# write_csv(lm_df, here('files', 'lm_df.csv'))
# 
# psm_df <- read_csv(here('files_main', 'psm_df.csv')) %>% 
#   mutate(set = ifelse(dataset <= 200, 'main', 'linear')) %>%
#   bind_rows(read_csv(here('files_small', 'psm_df.csv')) %>% mutate(set = 'small')) %>%
#   bind_rows(read_csv(here('files_large', 'psm_df.csv')) %>% mutate(set = 'large')) %>%
#   left_join(index_df)
# 
# write_csv(lm_df, here('files', 'psm_df.csv'))
# 
# 
# aipw_df <- read_csv(here('files_main/homemade', 'aipw.csv')) %>% 
#   mutate(set = ifelse(dataset <= 200, 'main', 'linear')) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'aipw.csv')) %>% mutate(set = 'small')) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'aipw.csv')) %>% mutate(set = 'large')) %>%
#   left_join(index_df)
# 
# write_csv(aipw_df, here('files/homemade', 'aipw.csv'))
# 
# 
# aipw_trunc_df <- read_csv(here('files_main/homemade', 'aipw_trunc.csv')) %>% 
#   mutate(set = ifelse(dataset <= 200, 'main', 'linear')) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'aipw_trunc.csv')) %>% mutate(set = 'small')) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'aipw_trunc.csv')) %>% mutate(set = 'large')) %>%
#   left_join(index_df)
# 
# write_csv(aipw_trunc_df, here('files/homemade', 'aipw_trunc.csv'))
# 
# tmle_df <- read_csv(here('files_main/homemade', 'tmle.csv')) %>% 
#   mutate(set = ifelse(dataset <= 200, 'main', 'linear')) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'tmle.csv')) %>% mutate(set = 'small')) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'tmle.csv')) %>% mutate(set = 'large')) %>%
#   left_join(index_df)
# 
# write_csv(tmle_df, here('files/homemade', 'tmle.csv'))
# 
# dml_ols_logit_df <- read_csv(here('files_main/homemade', 'dml_ols_logit.csv')) %>%
#   mutate(dataset = 1:nrow(.)) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'dml_ols_logit.csv')) %>% 
#               mutate(set = 'small',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'dml_ols_logit.csv')) %>% 
#               mutate(set = 'large',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   left_join(index_df) 
# 
# write_csv(dml_ols_logit_df, here('files/homemade', 'dml_ols_logit.csv'))
# 
# dml_grf_df <- read_csv(here('files_main/homemade', 'dml_grf.csv')) %>%
#   mutate(dataset = 1:nrow(.)) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'dml_grf.csv')) %>% 
#               mutate(set = 'small',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'dml_grf.csv')) %>% 
#               mutate(set = 'large',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   select(-dorie_dataset) %>%
#   left_join(index_df)
# 
# write_csv(dml_grf_df, here('files/homemade', 'dml_grf.csv'))
# 
# 
# dml_superlearner_df <- read_csv(here('files_main/homemade', 'dml_superlearner.csv')) %>%
#   mutate(dataset = 1:nrow(.)) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'dml_superlearner.csv')) %>% 
#               mutate(set = 'small',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'dml_superlearner.csv')) %>% 
#               mutate(set = 'large',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   select(-dorie_dataset) %>%
#   left_join(index_df)
# 
# write_csv(dml_superlearner_df, here('files/homemade', 'dml_superlearner.csv'))

# ols_logit_pred_combined <- readRDS(here('files_main/homemade', 'ols_logit_pred.RDS')) %>%
#   append(readRDS(here('files_small/homemade', 'ols_logit_pred.RDS'))) %>%
#   append(readRDS(here('files_large/homemade', 'ols_logit_pred.RDS')))
# 
# grf_pred_combined <- readRDS(here('files_main/homemade', 'grf_pred.RDS')) %>%
#   append(readRDS(here('files_small/homemade', 'grf_pred.RDS'))) %>%
#   append(readRDS(here('files_large/homemade', 'grf_pred.RDS')))
# 
# superlearner_pred_combined <- readRDS(here('files_main/homemade', 'superlearner_pred.RDS')) %>%
#   append(readRDS(here('files_small/homemade', 'superlearner_pred.RDS'))) %>%
#   append(readRDS(here('files_large/homemade', 'superlearner_pred.RDS')))
# 
# saveRDS(ols_logit_pred_combined, here('files/homemade', 'ols_logit_pred.RDS'))
# saveRDS(grf_pred_combined, here('files/homemade', 'grf_pred.RDS'))
# saveRDS(superlearner_pred_combined, here('files/homemade', 'superlearner_pred.RDS'))

# dml_ols_logit <- read_csv(here('files_large/homemade', 'dml_ols_logit.csv')) 
# 
# dml_ols_logit %>%
#   filter(set == 'main') %>%
#   write_csv(here('files_main/homemade', 'dml_ols_logit.csv'))
# 
# dml_ols_logit %>%
#   filter(set == 'large') %>%
#   write_csv(here('files_large/homemade', 'dml_ols_logit.csv'))

## add lalonde ####
# read_csv(here('files_combined', 'lm_df.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde', 'lm_df.csv'))) %>%
#   write_csv(here('files', 'lm_df.csv'))
# 
# 
# read_csv(here('files_combined', 'psm_df.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde', 'psm_df.csv'))) %>%
#   write_csv(here('files', 'psm_df.csv'))
# 
# read_csv(here('files_combined', 'lin.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde', 'lin.csv'))) %>%
#   write_csv(here('files', 'lin.csv'))
# 
# read_csv(here('files_combined', 'gcomp.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde', 'gcomp.csv'))) %>%
#   write_csv(here('files', 'gcomp.csv'))
# 
# read_csv(here('files_combined/homemade', 'aipw.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde/homemade', 'aipw.csv'))) %>%
#   write_csv(here('files/homemade', 'aipw.csv'))
# 
# read_csv(here('files_combined/homemade', 'aipw_trunc.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde/homemade', 'aipw_trunc.csv'))) %>%
#   write_csv(here('files/homemade', 'aipw_trunc.csv'))
# 
# read_csv(here('files_combined/homemade', 'tmle.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde/homemade', 'tmle.csv'))) %>%
#   write_csv(here('files/homemade', 'tmle.csv'))
# 
# read_csv(here('files_combined/homemade', 'dml_ols_logit.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde/homemade', 'dml_ols_logit.csv'))) %>%
#   write_csv(here('files/homemade', 'dml_ols_logit.csv'))
# 
# read_csv(here('files_combined/homemade', 'dml_grf.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde/homemade', 'dml_grf.csv'))) %>%
#   write_csv(here('files/homemade', 'dml_grf.csv'))
# 
# read_csv(here('files_combined/homemade', 'dml_superlearner.csv')) %>%
#   bind_rows(read_csv(here('files_lalonde/homemade', 'dml_superlearner.csv'))) %>%
#   write_csv(here('files/homemade', 'dml_superlearner.csv'))
# 
# 
# readRDS(here('files_combined/homemade', 'ols_logit_pred.RDS')) %>%
#   append(readRDS(here('files_lalonde/homemade', 'ols_logit_pred.RDS'))) %>%
#   saveRDS(here('files/homemade', 'ols_logit_pred.RDS'))
# 
# readRDS(here('files_combined/homemade', 'grf_pred.RDS')) %>%
#   append(readRDS(here('files_lalonde/homemade', 'grf_pred.RDS'))) %>%
#   saveRDS(here('files/homemade', 'grf_pred.RDS'))
# 
# readRDS(here('files_combined/homemade', 'superlearner_pred.RDS')) %>%
#   append(readRDS(here('files_lalonde/homemade', 'superlearner_pred.RDS'))) %>%
#   saveRDS(here('files/homemade', 'superlearner_pred.RDS'))

# dml_ols_logit_df <- read_csv(here('files_main/homemade', 'dml_ols_logit.csv')) %>%
#   mutate(dataset = 1:nrow(.)) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'dml_ols_logit.csv')) %>% 
#               mutate(set = 'small',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'dml_ols_logit.csv')) %>% 
#               mutate(set = 'large',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   left_join(index_df) 
# 
# write_csv(dml_ols_logit_df, here('files/homemade', 'dml_ols_logit.csv'))
# 
# dml_grf_df <- read_csv(here('files_main/homemade', 'dml_grf.csv')) %>%
#   mutate(dataset = 1:nrow(.)) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'dml_grf.csv')) %>% 
#               mutate(set = 'small',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'dml_grf.csv')) %>% 
#               mutate(set = 'large',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   select(-dorie_dataset) %>%
#   left_join(index_df)
# 
# write_csv(dml_grf_df, here('files/homemade', 'dml_grf.csv'))
# 
# 
# dml_superlearner_df <- read_csv(here('files_main/homemade', 'dml_superlearner.csv')) %>%
#   mutate(dataset = 1:nrow(.)) %>%
#   bind_rows(read_csv(here('files_small/homemade', 'dml_superlearner.csv')) %>% 
#               mutate(set = 'small',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   bind_rows(read_csv(here('files_large/homemade', 'dml_superlearner.csv')) %>% 
#               mutate(set = 'large',
#                      dorie_dataset = 7,
#                      dataset = row_number())) %>%
#   select(-dorie_dataset) %>%
#   left_join(index_df)
# 
# write_csv(dml_superlearner_df, here('files/homemade', 'dml_superlearner.csv'))

# ols_logit_pred_combined <- readRDS(here('files_main/homemade', 'ols_logit_pred.RDS')) %>%
#   append(readRDS(here('files_small/homemade', 'ols_logit_pred.RDS'))) %>%
#   append(readRDS(here('files_large/homemade', 'ols_logit_pred.RDS')))
# 
# grf_pred_combined <- readRDS(here('files_main/homemade', 'grf_pred.RDS')) %>%
#   append(readRDS(here('files_small/homemade', 'grf_pred.RDS'))) %>%
#   append(readRDS(here('files_large/homemade', 'grf_pred.RDS')))
# 
# superlearner_pred_combined <- readRDS(here('files_main/homemade', 'superlearner_pred.RDS')) %>%
#   append(readRDS(here('files_small/homemade', 'superlearner_pred.RDS'))) %>%
#   append(readRDS(here('files_large/homemade', 'superlearner_pred.RDS')))
# 
# saveRDS(ols_logit_pred_combined, here('files/homemade', 'ols_logit_pred.RDS'))
# saveRDS(grf_pred_combined, here('files/homemade', 'grf_pred.RDS'))
# saveRDS(superlearner_pred_combined, here('files/homemade', 'superlearner_pred.RDS'))

# dml_ols_logit <- read_csv(here('files_large/homemade', 'dml_ols_logit.csv')) 
# 
# dml_ols_logit %>%
#   filter(set == 'main') %>%
#   write_csv(here('files_main/homemade', 'dml_ols_logit.csv'))
# 
# dml_ols_logit %>%
#   filter(set == 'large') %>%
#   write_csv(here('files_large/homemade', 'dml_ols_logit.csv'))

# Introduction

In causal inference, functional form misspecification of underlying models can bias estimates of treatment effects [@hernan_2020_causal; @morgan_2015_counterfactuals]. There have been two important developments that attempt to overcome this. First, methodologists have developed machine learning methods that allow greater flexibility in estimation, adjusting for covariates in data-driven, complex ways [@balzer_2021_invited; @brand_2023_recent]. The second development is double robust methods [@bang_2005_doubly; @kang_2007_demystifying], which estimate two models: one for treatment exposure and another for the outcome. These models are robust to misspecification of either one of these "nuisance" models.

A number methods unifying these two developments have proliferated. These double robust methods for flexible covariate adjustment use machine learning methods to adapatively model the data generating processes at play. These models purportedly overcome the shortcomings of both traditional statistical methods and machine learning methods. Common statistical methods -- such as OLS regression and matching on propensity scores estimated from logistic regression -- have rigid functional form assumptions and fail to calculate stable estimates when the number of covariates is large relative to the number of observations. Machine learning methods, on the other hand, are often difficult to interpret. They can also suffer from overfitting, where the flexibility of the model becomes a weakness and predictions out-of-sample are poor, yet efforts to correct for overfitting can introduce regularization bias [@hastie_2009_elements]. Double robust methods with machine learning dispose of the constricting functional form assumptions of common statistical methods, and they correct for the regularization bias of flexible machine learning methods. They can also accommodate large numbers of covariates and produce easily interpretable treatment effect estimates. Despite their apparent advantages, these methods remain rarely utilized by social scientists. Part of the barrier has been lack of familiarity with these methods. It has also been unclear how these methods compare, or whether such methods actually perform better than traditional methods in finite samples.

This paper makes advances on these fronts. First, it is a guide to some of the latest methods in double robust, flexible covariate adjustment for causal inference, explaining the methods to a social scientist audience. Methods covered include Augmented Inverse Probability Weighting (AIPW), Targeted Maximum Likelihood Estimation (TMLE), and Double or Debiased Machine Learning (DML). This paper reviews the theory behind these methods as well as simple R implementations.

Second, this paper evaluates these methods. They are tested on simulations from @dorie_2019_automated, which cover a range of data-generating processes where ignorability holds -- i.e., there are no unmeasured confounders. They are then used to estimate experimental and observational effects from real-world data, the National Support for Work Demonstration (NSW) originally analyzed by @lalonde_1986_evaluating. In these evaluations, double robust methods are compared to "single robust" methods (i.e., ones with one nuisance model). These include traditional or simpler statistical methods commonly used by social scientists: ordinary least squares (OLS) regression, matching on propensity scores estimated from logistic regression (PSM), and inverse probability weighting (IPW). They are also compared to two more flexible methods that may overcome the misspecification issues that double robust methods aim to overcome: G-computation and the Lin estimator [@lin_2013_agnostic].

Results from simulations where ignorability holds show that some double robust methods outperform traditional statistical methods, but only slightly. AIPW and TMLE perform the best, at least when used in conjunction with flexible machine learning algorithms, while DML does slightly worse than traditional methods. G-computation with flexible machine learning performs as well as the lowest-error double robust methods. Despite their lower average error, the computation times of these flexible methods are high. With its still relatively low error and much faster computation time, OLS remains a sensible choice as an estimator, and the Lin estimator -- which is as quick as standard OLS regression -- performs only slightly worse than the best double robust methods.

Results from the NSW study highlight the importance of ignorability. When observational samples are not comparable to the experimental sample, or when important covariates are not included, all methods -- double robust or traditional -- fail to recover experimental estimates. When appropriate covariates are included, most methods perform well. 
<!-- Exceptions include IPW, DML, and the Lin estimator (surprising given the simulation results), whose estimates tend to be less stable across samples.   -->

# Motivation

## Literature Review

Although previous work has compared different methods for covariate adjustment, there has not yet been a thorough comparison and evaluation of the recent and popular double robust methods of AIPW, TMLE, and DML for a social scientist audience.

Although some introductions to double robust methods exist, they do not discuss them in the context of covariate adjustment for causal inference, or their treatment is overly technical for a social scientist audience. For example, @kang_2007_demystifying provide an an excellent overview and evaluation of double robust methods, but in the context of missing data, and the authors consider AIPW but not TMLE or DML. @bang_2005_doubly introduce double robust models for both causal inference and missing data, but their treatment is rather technical, and they only discuss AIPW. @lundberg_2022_researcher provide brief schematic overviews of double robust methods for a social scientist audience, but they do not evaluate these methods.

>>>>

---
output:
  # bookdown::word_document2:
    # reference_docx: "word-template.docx"
  bookdown::pdf_document2:
    toc: yes
    keep_tex: true
    number_sections: yes
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #latex_engine: xelatex
always_allow_html: true
header-includes:
  #- \usepackage{setspace}\doublespace
  # - \usepackage[nolists, fighead, tabhead]{endfloat}
  # - \usepackage{endnotes}
  # - \let\footnote=\endnote
  # - \setlength{\headheight}{14.5pt}
  # - \setlength{\headheight}{13.6pt}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \lhead{N.I. Hoffmann} 
  - \rhead{`r format(Sys.time(), '%B %e, %Y')`}
  - \counterwithin{figure}{section}
  - \counterwithin{table}{section}
# - \newcommand{\beginsupplement}{\setcounter{table}{0}  
# \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} 
# \renewcommand{\thefigure}{A\arabic{figure}}}

@@

title: "Online Appendix"
subtitle: "Double Robust, Flexible Adjustment Methods for Causal Inference: An Overview and an Evaluation"

@@

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = T, dev = c('pdf', 'png'), fig.retina = 3, ft.latex.float = 'float')
options("yaml.eval.expr" = TRUE)

@@

perform <- function(est_df, label){
  est_df %>%
    # mutate(d = d/truth,
    #        truth = 1) %>%
    summarize(bias = mean(d - truth),
              percent_bias = bias/sd(d),
              rmse = sqrt(mean((ate - truth)^2)),
              mae = median(abs(d - truth))
              # fail_count = first(fail_count),
              # comp_time = first(comp_time)
              ) %>%
    mutate(label = label,
           n = nrow(est_df) + fail_count) %>%
    select(label, everything()) %>%
    return()
}

model_matrix <- function(...){
  options(na.action='na.pass')
  matrix_out <- model.matrix(...)
  options(na.action = 'na.omit')
  return(matrix_out)
}

compare_tbl <- function(model_list, treatment){
  out_list <- list()
  k <- 1
  for(model_set in model_list){
    names(model_set[[2]]) <- c('estimate', 'std.error')

    out_list[[k]] <- tidy(model_set[[1]]) %>%
      filter(term == treatment) %>%
      bind_rows(bind_rows(model_set[[2]])) %>%
      bind_rows(model_set[[3]]) %>%
      mutate(term = c('Original', 'AIPW (GRF)', 'DML (SuperLearner)'), p.value = p_val(estimate, std.error))

    k <- k+1
  }
  return(out_list)
}

dml_grf <- function(outcome, treatment, covariates, clustervar = NULL, dataset, 
                    fe = T, target = 'treated', paper, model, drop_na_grf = F, seed = 123){
  set.seed(seed)

  dataset <- filter(dataset, !is.na(!!sym(outcome))) %>%
    select(all_of(c(treatment, outcome, covariates, clustervar)))

  if(drop_na_grf == T){
    dataset <- dataset %>%
      drop_na()
  }

  y <- as.numeric(dataset[[outcome]])
  X <- model_matrix(~., select(dataset, all_of(covariates)))[,]
  d <- as.numeric(dataset[[treatment]])

  grf_model <- grf::causal_forest(X = X, Y = y, W = d, seed = 123,
                        clusters = dataset[[clustervar]])

  grf_out <- grf::average_treatment_effect(grf_model, target.sample = target)

  dataset_drop_na <- as.data.frame(model_matrix(~., dataset)[,]) %>%
    # remove annoying characters
    rename_with(., ~ gsub("'", "", iconv(.x, from = "UTF-8", to='ASCII//TRANSLIT'))) %>%
    rename_with(., ~ gsub("\\[|\\]|\\/|\\*|\\)|\\(", "", .x)) %>%
    rename_with(., ~ gsub(" ", "", .x)) %>%
    # drop missing
    drop_na() %>%
    # remove uninformative columns
    select(where(~n_distinct(.) > 1)) 
  dataset_drop_na <- dataset_drop_na %>%
    left_join(dataset_drop_na %>%
                group_by(!!sym(clustervar)) %>%
                summarize(across(everything(), mean, .names = 'mean_{.col}')) %>%
                ungroup()) %>%
    mutate(intercept = 1)
  # covariates_design <- dataset_drop_na %>% 
  #   select(select(-c(treatment, outcome, clustervar))) %>%
  #   names() %>%
  #   append('intercept', after = 0)
  covariates_design <- names(dataset_drop_na)[!(names(dataset_drop_na) %in% c(treatment, outcome, clustervar)) &
                                                !str_detect(names(dataset_drop_na),  'mean_')]

  # dataset_drop_na$cov_mean <- dataset_drop_na %>%
  #   select(all_of(covariates_design)) %>%
  #   mutate_all(scale) %>%
  #   apply(1, mean)

  # dml_dataframe <- as.data.frame(cbind(y, d, 
  #                                      d_bar = dataset_drop_na[[paste0('mean_', treatment)]],
  #                                      as_tibble(X)[,-1], 
  #                                      as_tibble(mean_X)[,-1], 
  #                                      cluster = dataset_drop_na[[clustervar]]))

  graph_ensemble_regr = gunion(list(
      po("learner", lrn("regr.cv_glmnet", s = "lambda.min")),
      po("learner", lrn('regr.xgboost', max_depth = 4)),
      po("learner", lrn("regr.glm"))
    )) %>>%
      po("regravg", 3)

  ensemble_pipe_regr = as_learner(graph_ensemble_regr)

  # DML
  set.seed(seed)
  dml_data <- double_ml_data_from_data_frame(dataset_drop_na,
                                             x_cols = covariates_design,
                                             y_col = outcome,
                                             d_cols = treatment,
                                             cluster_cols = clustervar)

  # y <- as.numeric(dataset_drop_na[[outcome]])
  # # X <- select(dataset_drop_na, covariates_design)
  # X <- model_matrix(~., select(dataset_drop_na, all_of(covariates)))[,]
  # # mean_X <- model_matrix(~., select(dataset_drop_na, all_of(paste0('mean_', covariates))))[,]
  # d <- as.numeric(dataset_drop_na[[treatment]])
  # 
  # dml_data <- double_ml_data_from_matrix(X = X, y = y, d = d, 
  #                                        cluster_vars = dataset_drop_na[[clustervar]])

  # dml_data <- double_ml_data_from_matrix(X = select(dataset_drop_na, all_of(covariates_design)),
  #                                            y = dataset_drop_na[,outcome],
  #                                            d = dataset_drop_na[,treatment],
  #                                            cluster_vars = dataset_drop_na[,clustervar])
  obj_dml_plr_sim_pipe_ensemble = DoubleMLPLR$new(dml_data,
                                                  ml_l = ensemble_pipe_regr,
                                                  ml_m = ensemble_pipe_regr)
  obj_dml_plr_sim_pipe_ensemble$fit() 
  dml_out <- data.frame(estimate = obj_dml_plr_sim_pipe_ensemble$coef,
                         std.error = obj_dml_plr_sim_pipe_ensemble$se)

  if(fe == T){
    # DML CRE
    set.seed(seed)  
    dml_data_cre <- XTDML::dml_cre_data_from_data_frame(dataset_drop_na,
                                                 x_cols = covariates_design,
                                                 y_col = outcome,
                                                 d_cols = treatment,
                                                 xbar_cols = paste0('mean_', covariates_design[-length(covariates_design)]),
                                                 dbar_cols = paste0('mean_', treatment),
                                                 cluster_cols = clustervar)  
    obj_dml_cre = XTDML::dml_cre_plr$new(dml_data_cre,
                                         ml_l = ensemble_pipe_regr,
                                         ml_m = ensemble_pipe_regr)

    obj_dml_cre$fit()
    dml_cre_out <- data.frame(estimate = obj_dml_cre$coef_theta,
                           std.error = obj_dml_cre$se_theta)

  # DML WG-CRE
  require(parameters)
  set.seed(seed)  
  dml_data_wg_cre <- XTDML::dml_hybrid_data_from_data_frame(dataset_drop_na,
                                               x_cols = covariates_design,
                                               y_col = outcome,
                                               d_cols = treatment,
                                               xbar_cols = paste0('mean_', covariates_design[-length(covariates_design)]),
                                               dbar_cols = paste0('mean_', treatment),
                                               cluster_cols = clustervar)  
  obj_dml_wg_cre = XTDML::dml_hybrid_plr$new(dml_data_wg_cre,
                                       ml_l = ensemble_pipe_regr,
                                       ml_m = ensemble_pipe_regr)

  obj_dml_wg_cre$fit()
  dml_wg_cre_out <- data.frame(estimate = obj_dml_wg_cre$coef_theta,
                         std.error = obj_dml_wg_cre$se_theta)
  } else {
    dml_cre_out <- data.frame(estimate = NA,
                           std.error = NA)
    dml_wg_cre_out <- data.frame(estimate = NA,
                         std.error = NA)
  }

  return(list(grf = grf_out, 
              dml = dml_out,
              dml_cre = dml_cre_out,
              dml_wg_cre = dml_wg_cre_out))
}
```
